What is essential? -- A pilot survey on views about the requirements metamodel of reqT

### 2015-10-01 paper/paper-refsq2016-submission.tex
update paper

% first the title is needed
\title{reqT.org --  Towards a Semi-Formal, Open and Scalable Requirements Modeling Tool}
%reqT --  A Requirements Modeling Tool for Code Lovers
% a short form should be given in case it is too long for the running head
%\titlerunning{DRAFT MANUSCRIPT - Please do not cite. Submitted to REFSQ'13.}

@@

%%%% ABSTRACT
\begin{abstract}
[{\bf Context and motivation}] This research preview presents ongoing work on a free software requirements modeling tool called reqT that is developed in an educational context. [{\bf Question/problem}] The work aims to engage computer science students in Requirements Engineering (RE) through a tool that captures essential RE concepts in executable code. [{\bf Principal ideas}] Requirements are modeled using an internal DSL in the Scala programming language that blends natural language strings with a graph-oriented formalism. [{\bf Contribution}] The metamodel of reqT and its main features are presented and modeling examples are provided together with a discussion on initial experiences from student projects, limitations and directions of further research.   
\keywords{requirements engineering, requirements modeling, software engineering, CASE tool, requirements metamodel, requirements engineering education, internal DSL, embedded DSL, Scala programming language}
\end{abstract}

>>>>

% first the title is needed
\title{What is essential? -- A pilot survey on views about the requirements metamodel of reqT}
%reqT --  A Requirements Modeling Tool for Code Lovers
% a short form should be given in case it is too long for the running head
%\titlerunning{DRAFT MANUSCRIPT - Please do not cite. Submitted to REFSQ'13.}

@@

%%%% ABSTRACT
\begin{abstract}
[{\bf Context and motivation}] This research preview presents ongoing work on the metamodel of a free software requirements modeling tool called reqT that is developed in an educational context. [{\bf Question/problem}] The work aims to make an initial validation of a survey instrument that elicits views on the metamodel of the reqT tool, which seek to engage computer science students in Requirements Engineering (RE) through an open source requirements engineering DSL embedded in the Scala programming language. The research question underpinning the presented work is: Which RE concepts are essential to include in the metamodel for a requirements engineering tool in an educational context?  [{\bf Principal ideas}] A survey instrument is developed with a list of 92 concepts (49 entities, 15 relations and 28 attributes) and a set of questions for each concept that elicit the respondents' views on the usage and interpretation of each concept.  [{\bf Contribution}] The survey is initially validated in a pilot study involving 14 Swedish RE scholars as subjects. The survey results indicate that the survey is feasible if the respondents is willing to invest around 30 minutes of their time. The analysis of the responses suggest that many of the concepts in the metamodel are used frequently by the respondents and there is a large degree of agreement among the respondents about the meaning of the concepts. The results are encouraging for future work on empirical validation of the relevance of the reqT metamodel.

\keywords{requirements engineering, requirements metamodel, CASE tool, requirements engineering education, embedded domain-specific language, empirical software engineering}
\end{abstract}
